---
title: "Microdialysis 16S analysis"
author: "Andreas Schneider"
date: "05/07/2023"
output: 
  html_document:
    code_folding: hide
    toc: true
    toc_float: true
    number_sections: true
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
suppressMessages({
  library(here)
  library(vegan)
  library(RColorBrewer)
  library(ggplot2)
  library(gridExtra)
  library(phyloseq)
  library(reshape2)
  library(lefser)
  library(zinbwave)
  library(pheatmap)
  library(DESeq2)
  library(agricolae)
  library(grid)
  library(gridExtra)
  library(dplyr)
  source(here("UPSCb-common/src/R/featureSelection.R"))
  source(here("Rtoolbox/src/utilsDE.r"))
})

```



# Import Data

```{r}
taxa <- read.delim(here("data/NO_PLANT_ASV_tax_species.txt"))
colnames(taxa)[1] <- "SHA_ID"
taxa$ASV_ID <- paste0("ASV_", 1:nrow(taxa))
rownames(taxa) <- taxa$ASV_ID

#Clean up taxonomy.
taxa$Species <- ifelse(taxa$Species_exact!="", taxa$Species_exact, taxa$Species)


taxa[taxa==""] <- NA


taxa[is.na(taxa)] <- "unidentified"

taxa$Phylum <- ifelse(grepl("unidentified", taxa$Phylum), 
                         paste0("Unclassified.", taxa$Kingdom),
                         taxa$Phylum)

#Let's make a function to automate this
fixNAtax <- function(tax, rank) {
  coln <- which(colnames(tax)==rank)
  namen <- colnames(tax)[coln]
  namen1 <- colnames(tax)[coln-1]
  tax[,namen] <- ifelse(grepl("Unclassified", tax[,namen1]),
                        tax[,namen1],
                        tax[,namen])
  tax[,namen] <- ifelse(grepl("unidentified", tax[,namen]),
                        paste0("Unclassified.", tax[,namen1]),
                        tax[,namen])
  return(tax)
}
taxa2 <- fixNAtax(taxa, "Class")
taxa2 <- fixNAtax(taxa2, "Order")
taxa2 <- fixNAtax(taxa2, "Family")
taxa2 <- fixNAtax(taxa2, "Genus")
taxa2 <- fixNAtax(taxa2, "Species")

taxa2$Species <- ifelse(grepl("Unclassified", taxa2$Species), taxa2$Species, paste0(taxa2$Genus, "_", taxa2$Species))

#Import counts
seqtab <- read.delim(here("data/NO_PLANT_ASV_table.txt"), row.names = 1)
seqtab <- seqtab[,-c(15)]

all(rownames(seqtab)==taxa$SHA_ID)
rownames(seqtab) <- taxa2$ASV_ID

meta <- read.delim(here("doc/metadata_Z.txt"), row.names = NULL)
colnames(meta)[1:3] <- colnames(meta)[2:4]
#remove color column; remove mock
meta <- meta[-18,-c(3,4)]

meta$perfusate <- ifelse(grepl("MQ", meta$condition), "MQ", "Sucrose")
meta$Litter <- ifelse(grepl("L", meta$condition), "Litter", "No_Litter")
meta$condition2 <- paste0(meta$perfusate, ".", meta$Litter)
```

# Read counts

```{r}
boxplot(colSums(seqtab)~meta$condition2)

```

Better than in the ITS dataset, checking optimal cycle number is worth it.

# ASVs unique to Sucrose/MQ (ignoring litter)

```{r}
seqtab <- seqtab[,meta$ID]

##Filtering low abundance ASVs
ccc <- as.factor(meta$perfusate)
names(ccc) <- meta$ID
#dim(seqtab)
seqtab.f1 <- seqtab[featureSelect(data.matrix(seqtab), ccc, 3, 2),]
#dim(seqtab.f1)
seqtab.f2 <- seqtab.f1[featureSelectProp(data.matrix(seqtab.f1), ccc, 0.00005),]
dim(seqtab.f2)

##Overlap between sucrose and mq
seqtab_MQ <- seqtab.f2[,meta$condition2=="MQ.No_Litter"]
seqtab_MQ <- seqtab_MQ[rowSums(seqtab_MQ)>sum(seqtab_MQ)*0.00003,]

seqtab_su <- seqtab.f2[,meta$condition2=="Sucrose.No_Litter"]
seqtab_su <- seqtab_su[rowSums(seqtab_su)>sum(seqtab_su)*0.00003,]

length(intersect(rownames(seqtab_MQ), rownames(seqtab_su)))
length(setdiff(rownames(seqtab_MQ), rownames(seqtab_su)))
length(setdiff(rownames(seqtab_su), rownames(seqtab_MQ)))
```

Similar to ITS, more unique ASVs in MQ samples.

## ASVs unique between other comparisons

### MQ vs MQ + litter

```{r}
seqtab_MQ.L <- seqtab.f2[,meta$condition2=="MQ.Litter"]
seqtab_MQ.L <- seqtab_MQ.L[rowSums(seqtab_MQ.L)>sum(seqtab_MQ.L)*0.00003,]

length(intersect(rownames(seqtab_MQ), rownames(seqtab_MQ.L)))
length(setdiff(rownames(seqtab_MQ), rownames(seqtab_MQ.L)))
length(setdiff(rownames(seqtab_MQ.L), rownames(seqtab_MQ)))

```

### MQ+litter vs Sucrose+litter

```{r}
seqtab_su.L <- seqtab.f2[,meta$condition2=="Sucrose.Litter"]
seqtab_su.L <- seqtab_su.L[rowSums(seqtab_su.L)>sum(seqtab_su.L)*0.00003,]

length(intersect(rownames(seqtab_MQ.L), rownames(seqtab_su.L)))
length(setdiff(rownames(seqtab_MQ.L), rownames(seqtab_su.L)))
length(setdiff(rownames(seqtab_su.L), rownames(seqtab_MQ.L)))
```

## ASVs common for all treatments

```{r}
asvs_common_all <- Reduce(intersect, list(rownames(seqtab_MQ), 
                       rownames(seqtab_su), 
                       rownames(seqtab_MQ.L), 
                       rownames(seqtab_su.L)))
length(asvs_common_all)

taxa_com_all <- taxa2[asvs_common_all,c(4:9,13)]

```

## ASVs unique to only one of the treatments

Order: MQ, MQ+litter, Sucrose, Sucrose+litter

```{r}
asvs_mq <- rownames(seqtab_MQ)
asvs_mq.l <- rownames(seqtab_MQ.L)
asvs_su <- rownames(seqtab_su)
asvs_su.l <- rownames(seqtab_su.L)

#Unique to MQ
sum(!asvs_mq%in%unique(c(asvs_mq.l, asvs_su, asvs_su.l)))
#Unique to MQ Litter
sum(!asvs_mq.l%in%unique(c(asvs_mq, asvs_su, asvs_su.l)))
#Unique to Sucrose
sum(!asvs_su%in%unique(c(asvs_mq.l, asvs_mq, asvs_su.l)))
#Unique to Sucrose+litter
sum(!asvs_su.l%in%unique(c(asvs_mq.l, asvs_su, asvs_mq)))

```
```{r}
#MQ
mq_ex <- asvs_mq[!asvs_mq%in%unique(c(asvs_mq.l, asvs_su, asvs_su.l))]

seqtab_MQ_p <- prop.table(data.matrix(seqtab_MQ), margin = 2)*100
seqtab_MQ_p_ex <- seqtab_MQ_p[mq_ex,]

mq_ex_tab <- cbind(Mean=rowMeans(seqtab_MQ_p_ex),
                   Presence=rowSums(seqtab_MQ_p_ex>0),
                   taxa2[mq_ex,c(4:9)])

#MQ+Litter
mq_ex_l <- asvs_mq.l[!asvs_mq.l%in%unique(c(asvs_mq, asvs_su, asvs_su.l))]

seqtab_MQ.L_p <- prop.table(data.matrix(seqtab_MQ.L), margin = 2)*100
seqtab_MQ.L_p_ex <- seqtab_MQ.L_p[mq_ex_l,]

mq_ex_l_tab <- cbind(Mean=rowMeans(seqtab_MQ.L_p_ex),
                   Presence=rowSums(seqtab_MQ.L_p_ex>0),
                   taxa2[mq_ex_l,c(4:9)])

#Sucrose
su_ex <- asvs_su[!asvs_su%in%unique(c(asvs_mq.l, asvs_mq, asvs_su.l))]

seqtab_su_p <- prop.table(data.matrix(seqtab_su), margin = 2)*100
seqtab_su_p_ex <- seqtab_su_p[su_ex,]

su_ex_tab <- cbind(Mean=rowMeans(seqtab_su_p_ex),
                   Presence=rowSums(seqtab_su_p_ex>0),
                   taxa2[su_ex,c(4:9)])

#Sucrose+Litter
su_l_ex <- asvs_su.l[!asvs_su.l%in%unique(c(asvs_mq.l, asvs_su, asvs_mq))]

seqtab_su.L_p <- prop.table(data.matrix(seqtab_su.L), margin = 2)*100
seqtab_su.L_p_ex <- seqtab_su.L_p[su_l_ex,]

su_l_ex_tab <- cbind(Mean=rowMeans(seqtab_su.L_p_ex),
                   Presence=rowSums(seqtab_su.L_p_ex>0),
                   taxa2[su_l_ex,c(4:9)])

# mq_ex_tab
# su_ex_tab

ex_tab_all <- cbind(treatment_spec=c(rep("MQ", nrow(mq_ex_tab)), rep("MQ+Litter", nrow(mq_ex_l_tab)), rep("Sucrose", nrow(su_ex_tab)), rep("Sucrose+Litter", nrow(su_l_ex_tab))), rbind(mq_ex_tab, mq_ex_l_tab, su_ex_tab, su_l_ex_tab))
ex_tab_all2 <- ex_tab_all
ex_tab_all2$ASV <- rownames(ex_tab_all2)
ex_tab_all2 <- ex_tab_all2[,c(1, 4:9, 3, 10, 2)]
#write.csv(ex_tab_all2, here("results/analysis/uniq_sotus.csv"), row.names = TRUE, quote = FALSE)

```

# Rarefaction curves

```{r}
rarecurve(t(seqtab.f2), step = 50, label = FALSE, col = brewer.pal(3, "Dark2")[as.numeric(as.factor(meta$perfusate))])
abline(v = 40000, col = "red")
```

# Alpha diversity

## Richness (and rarefaction)

```{r}
#Rarefaction
set.seed(500)
seqtab.r <- t(rrarefy(t(seqtab.f2), min(colSums(seqtab.f2))))

seqtab.r_MQ <- t(rrarefy(t(seqtab_MQ), min(colSums(seqtab.f2))))
seqtab.r_su <- t(rrarefy(t(seqtab_su), min(colSums(seqtab.f2))))

rich_MQ <- colSums(seqtab.r_MQ>1)
rich_su <- colSums(seqtab.r_su>1)

boxplot(rich_MQ, rich_su, names=c("MQ", "Sucrose"))
wilcox.test(rich_MQ, rich_su, exact = FALSE)

```

Not significant.


### Richness including litter

```{r}
seqtab_MQ.L <- seqtab.f2[,meta$condition2=="MQ.Litter"]
seqtab_MQ.L <- seqtab_MQ.L[rowSums(seqtab_MQ.L)>sum(seqtab_MQ.L)*0.00003,]

seqtab_su.L <- seqtab.f2[,meta$condition2=="Sucrose.Litter"]
seqtab_su.L <- seqtab_su.L[rowSums(seqtab_su.L)>sum(seqtab_su.L)*0.00003,]

seqtab.r_MQ.L <- t(rrarefy(t(seqtab_MQ.L), min(colSums(seqtab.f2))))
seqtab.r_su.L <- t(rrarefy(t(seqtab_su.L), min(colSums(seqtab.f2))))

rich_MQ.L <- colSums(seqtab.r_MQ.L>1)
rich_su.L <- colSums(seqtab.r_su.L>1)

boxplot(rich_MQ, rich_su, rich_MQ.L, rich_su.L, names=c("MQ", "Sucrose", "MQ+Litter", "Sucrose+Litter"))
with(data.frame(Richness=c(rich_MQ, rich_su, rich_MQ.L, rich_su.L), 
                condition=meta$condition2[match(c(names(rich_MQ), names(rich_MQ.L), names(rich_su), names(rich_su.L)), meta$ID)]), kruskal(Richness, condition, console = TRUE))


#Same with ggplot
rich_nl <- data.frame(Richness=c(rich_MQ, rich_su, rich_MQ.L, rich_su.L), meta)
rich_nl$Litter <- factor(rich_nl$Litter, levels = c("No_Litter", "Litter"))

ggplot(rich_nl, aes(x = Litter, y = Richness))+
  geom_boxplot()+
  facet_wrap(~perfusate)+
  theme_classic()

ggsave("FigureS1D.pdf",
       device = "pdf",
       path = "results/figures",
       width = 116,
       height = 70,
       units = "mm")
```


## Shannon diversity

```{r}
divs <- as.data.frame(cbind(Shannon=diversity(t(seqtab.f2)), Simpson=diversity(t(seqtab.f2), index = "simpson")))
divs$perfusate <- meta$perfusate
divs$litter <- factor(meta$Litter, levels = c("No_Litter", "Litter"))
divs$condition <- factor(meta$condition2)

(shan1 <- ggplot(divs, aes(x = litter, y = Shannon))+
  geom_boxplot()+
  facet_wrap(~perfusate)+
  theme_minimal())

dir.create(file.path(here("results/figures")), recursive = TRUE, showWarnings = FALSE)

# ggsave("FigureS1B.pdf",
#        shan1,
#        device = "pdf",
#        path = "results/figures",
#        width = 116,
#        height = 70,
#        units = "mm")

div_krus <- with(divs, kruskal(Shannon, condition, console = TRUE))

divs_nl <- divs[divs$litter=="No_Litter",]
ggplot(divs_nl, aes(x = perfusate, y = Shannon))+
  geom_boxplot()+
  ggtitle("No Litter")+
  theme_minimal()

wilcox.test(Shannon~perfusate, data = divs[divs$litter=="No_Litter",])
```

# Beta Diversity

```{r}
taxa_ps <- as.matrix(taxa2[rownames(seqtab.r),])
rownames(meta) <- meta$ID
ps_rar <- phyloseq(otu_table(seqtab.r, taxa_are_rows = TRUE),
                   sample_data(meta),
                   tax_table(taxa_ps))
r.ord <- ordinate(ps_rar, "MDS", "bray")

#Scree Plot
percentages <- data.frame(PCo=paste0("PCo", seq(1, length(r.ord$values$Relative_eig), 1)),
                          percent=round(r.ord$values$Relative_eig*100, digits = 1))
barplot(percentages$percent, names.arg = percentages$PCo)

#PCoA plot

(ordplot1 <- plot_ordination(ps_rar, r.ord, type = "samples", color = "perfusate", shape = "Litter")+
    geom_point(size = 3)+
    xlab(paste0("PCoA 1 [", round(r.ord$values$Relative_eig[1]*100, digits = 1), "%]"))+
    ylab(paste0("PCoA 2 [", round(r.ord$values$Relative_eig[2]*100, digits = 1), "%]"))+
    theme(legend.position = "none")+
    theme_classic()
)
adonis2(phyloseq::distance(ps_rar, method = "bray")~condition2, data = meta)
adonis2(phyloseq::distance(ps_rar, method = "bray")~Litter, data = meta)
adonis2(phyloseq::distance(ps_rar, method = "bray")~perfusate, data = meta)

adonis2(t(seqtab.r)~perfusate*Litter, data = meta, by = "terms")

ggsave("Figure4B.pdf",
       ordplot1,
       device = "pdf",
       path = "results/figures",
       width = 100,
       height = 70,
       units = "mm")
```

Perfusate, litter, and the interaction between the two had a significant effect on community composition.

# Taxonomic composition

## Stacked bar charts

```{r}
#Barplot Phylum
count_melt <- melt(seqtab.r, varnames = c("ASV_ID", "ID"), value.name = "Count")
count_melt <- count_melt[count_melt$Count>0,]
count_melt <- merge.data.frame(count_melt, meta, by = "ID")
count_melt <- merge.data.frame(count_melt, taxa2, by = "ASV_ID")
#count_melt$SampleID <- factor(count_melt$SampleID, levels = c("MQ1", "MQ2","MQ4", "LMQ1", "LMQ2", "LMQ3", "LMQ4", "LMQ5", "S1", "S3", "S4", "S5", "LS1", "LS2", "LS3", "LS4", "LS5"))


```


## Stacked bar chart family - Phylum and Family

```{r}
#Create summarised matrices
count_melt2 <- melt(seqtab.r, varnames = c("ASV_ID", "ID"), value.name = "Count")
count_melt2 <- count_melt2[count_melt2$Count>0,]
count_melt2 <- merge.data.frame(count_melt2, meta, by = "ID")
count_melt2 <- merge.data.frame(count_melt2, taxa2, by = "ASV_ID")
class_mat <- acast(count_melt2, Class~ID, value.var = "Count", fun.aggregate = sum)
fam_mat <- acast(count_melt2, Family~ID, value.var = "Count", fun.aggregate = sum)
phylum_mat <- acast(count_melt2, Phylum~ID, value.var = "Count", fun.aggregate = sum)
```

```{r}
count_melt2_phyl <- count_melt2[,c(2,3,11)]
count_melt2_phyl2 <- count_melt2_phyl %>% 
  group_by(ID, Phylum) %>% 
  summarise(Count = sum(Count))

(pp2 <- ggplot(count_melt2_phyl2, aes(x = ID, y = Count, fill = Phylum))+
    geom_bar(stat = "identity", position = "fill")+
    scale_y_continuous(labels = scales::percent)+
    scale_fill_manual(values = c("#c176af",
                                "#74b93a",
"#b357c3",
"#58c173",
"#d24382",
"#629342",
"#6969ca",
"#bab049",
"#6c93cf",
"#dd9334",
"#4bc3b7",
"#d34e35",
"#3f8b64",
"#c25e65",
"#75722c",
"#b77746"))+
    ggtitle("Bar plot, Phylum level")+
    theme_classic()+
    #scale_fill_manual(values = sample(hue_pal()(length(unique(count_melt$Family)))))+
    theme(axis.text = element_text(angle = 45)))
```


```{r}
#### Bar plots with only top 15
fam_top15 <- rownames(fam_mat)[order(rowSums(fam_mat), decreasing = TRUE)][1:15]
fam_others <- rownames(fam_mat)[order(rowSums(fam_mat), decreasing = TRUE)][16:length(rownames(fam_mat))]

count_melt2 <- count_melt
count_melt2$Family2 <- ifelse(count_melt2$Family%in%fam_others, "Others", count_melt2$Family)
count_melt2$Family2 <- factor(count_melt2$Family2, levels = rev(c(fam_top15, "Others")))

count_melt2$ID <- factor(count_melt2$ID, levels = c("ZMQ1", "ZMQ2", "ZMQ4", 
                                                    "LMQ1", "LMQ2", "LMQ3", "LMQ4", "LMQ5",
                                                    "Z1", "Z3", "Z4", "Z5",
                                                    "LS1", "LS2", "LS3", "LS4", "LS5"))



count_melt2_fam <- count_melt2[,c(2,3,20)]
count_melt2_fam2 <- count_melt2_fam %>% 
  group_by(ID, Family2) %>% 
  summarise(Count = sum(Count))

(pf2 <- ggplot(count_melt2_fam2, aes(x = ID, y = Count, fill = Family2))+
    geom_bar(stat = "identity", position = "fill")+
    scale_y_continuous(labels = scales::percent)+
    scale_fill_manual(values = c("#d3d3d3",
                                 "#00b144",
                                 "#a673ff",
                                 "#b6d165",
                                 "#4094ff",
                                 "#908700",
                                 "#0192c1",
                                 "#b46b00",
                                 "#02dbd3",
                                 "#a11b2f",
                                 "#315e07",
                                 "#912b5c",
                                 "#ddc65b",
                                 "#ff8c85",
                                 "#89854f",
                                 "#feb68e"))+
    ggtitle("Bar plot, Family level Top 15")+
    theme_classic()+
    #scale_fill_manual(values = sample(hue_pal()(length(unique(count_melt$Family)))))+
    theme(axis.text = element_text(angle = 45)))
```

Print the figures

```{r}
# g_legend <- function(a.gplot){ 
#     tmp <- ggplot_gtable(ggplot_build(a.gplot)) 
#     leg <- which(sapply(tmp$grobs, function(x) x$name) == "guide-box") 
#     legend <- tmp$grobs[[leg]] 
#     legend
# } 
# 
# pp2_legend <- g_legend(pp2) 
# pf2_legend <- g_legend(pf2) 
# 
# pdf(here("results/figures/FigureS3B_legend.pdf"),
#     width = 6,
#     height = 8)
# grid.newpage()
# grid.draw(pp2_legend) 
# dev.off()
# 
# 
# 
# pdf(here("results/figures/Figure5B_legend.pdf"),
#     width = 10,
#     height = 8)
# grid.newpage()
# grid.draw(pf2_legend) 
# dev.off()
```

```{r}
#  ggsave("FigureS3B.pdf",
#    pp2 + theme(legend.position = "none"),
#        device = "pdf",
#        path = "results/figures",
#        width = 190,
#        height = 110,
#        units = "mm")
# 
# ggsave("Figure5B.pdf",
#    pf2 + theme(legend.position = "none"),
#        device = "pdf",
#        path = "results/figures",
#        width = 190,
#        height = 110,
#        units = "mm")

```

## Phylum level significant differences?

```{r}
set.seed(600)
meta2 <- meta
meta2$Litter <- as.factor(meta2$Litter)

sumobj_class <- SummarizedExperiment(assays <- list(as.matrix(phylum_mat)),
                               colData = meta2)

res_lefs_c <- lefser(sumobj_class, groupCol = "Litter")

lefserPlot(res_lefs_c)+
  theme_minimal()+
  ggtitle("Phylum level LDA scores")
```

0: Litter 1: No Litter

Proteobacteria higher with Litter, Bdellovibrionota without.

## Heatmap

```{r}
#HEATMAP
## Create summarised experiment object
meta$condition2 <- factor(meta$condition2, levels = c("MQ.No_Litter",
                                                      "MQ.Litter",
                                                      "Sucrose.No_Litter",
                                                      "Sucrose.Litter"))

sumobj <- SummarizedExperiment(assays <- list(as.matrix(seqtab.r)),
                               colData = meta)

sumobj_zinb <- zinbwave(sumobj, K = 2, epsilon = nrow(seqtab.r), normalizedValues = TRUE, residuals = TRUE)

seqtab.r_norm <- assay(sumobj_zinb, "normalizedValues")

sample_cols <- data.frame(Perfusate=meta$perfusate,Litter=meta$Litter, row.names = meta$ID)

sample_rows <- data.frame(Phylum=taxa2$Phylum[match(rownames(seqtab.r_norm), taxa2$ASV_ID)], row.names = taxa2$ASV_ID[match(rownames(seqtab.r_norm), taxa2$ASV_ID)])

ann_colors <- list(
  Litter = c(No_Litter="darkgreen", Litter="darkred")
)

pheatmap(seqtab.r_norm,
         show_rownames = FALSE,
         labels_col = meta$SampleID,
         #scale = "row",
         clustering_distance_cols = "correlation",
         annotation_col = sample_cols, 
         annotation_row = sample_rows,
         annotation_colors = ann_colors,
         clustering_method = "average")

pheatmap(seqtab.r_norm,
         show_rownames = FALSE,
         labels_col = meta$SampleID,
         #scale = "row",
         clustering_distance_cols = "correlation",
         annotation_col = sample_cols, 
         annotation_row = sample_rows,
         annotation_colors = ann_colors,
         filename = here("results/figures/FigureS4B.pdf"))
```

# Differential abundance

## MQ vs sucrose (excluding litter samples)

```{r}
sumobj_zinb2 <- zinbwave(sumobj, K = 0, epsilon = 1e12, residuals = TRUE, observationalWeights = TRUE)

#### Differential abundance, only pure mq and sucrose samples
dds_c2 <- DESeqDataSet(sumobj_zinb2, design = ~ condition2)
dds_c2 <- DESeq(dds_c2, sfType = "poscounts", useT = TRUE, minmu = 1e-6)

res_c2 <- results(dds_c2, contrast = c("condition2", "MQ.No_Litter", "Sucrose.No_Litter"))

res_c2_2 <- filterDE(res_c2)

da_asvs2 <- cbind(as.data.frame(res_c2_2)[,c(1,2,6)],taxa2[rownames(res_c2_2),-c(1:3)])

#PLOT DA ASVs
seqtab.r_p <- prop.table(seqtab.r, margin = 2)*100

seqtab.r_p_da <- seqtab.r_p[rownames(da_asvs2),]
seqtab.r_p_da <- seqtab.r_p_da[,!grepl("L", colnames(seqtab.r_p_da))]
da_perc <- reshape2::melt(seqtab.r_p_da, 
                          varnames = c("ASV_ID", "ID"), 
                          value.name = "Percentage")

da_perc <- merge.data.frame(da_perc, meta, by = "ID")
da_perc$ASV_Species <- paste0(da_perc$ASV_ID, "_", da_asvs2$Species)
da_perc$ASV_Species <- factor(da_perc$ASV_Species, levels = rev(unique(da_perc$ASV_Species)))

#da_perc 
asvs_sorted <- rownames(da_asvs2)[order(da_asvs2$log2FoldChange, decreasing = TRUE)]
asvs_sorted2 <- paste0(asvs_sorted, "_", da_asvs2$Species[order(da_asvs2$log2FoldChange, decreasing = TRUE)])
da_perc$ASV_Species <- factor(da_perc$ASV_Species, levels = asvs_sorted2)

ggplot(da_perc, aes(x = ASV_Species, y = Percentage, fill = perfusate))+
  geom_boxplot()+
  theme_minimal()+
  theme(axis.text.x = element_text(angle = 90))+
  coord_flip()+
  scale_y_continuous(trans='sqrt')

# ggsave("FigureS5B.pdf",
#        device = "pdf",
#        path = "results/figures",
#        width = 190,
#        height = 150,
#        units = "mm")
```
```{r}
#Table for paper
da_asvs2_cl <- da_asvs2[,c(4:9,13,2,3,1)]
da_asvs2_cl <- da_asvs2_cl[order(da_asvs2_cl$log2FoldChange),]
#Reverse lFC for table in paper.
da_asvs2_cl$log2FoldChange <- da_asvs2_cl$log2FoldChange*-1

#
```


## MQ vs MQ + Litter

```{r}
######### MQ vs MQ litter
res_mq_lit <- results(dds_c2, contrast = c("condition2", "MQ.No_Litter", "MQ.Litter"))

res_mq_lit2 <- filterDE(res_mq_lit)

da_asvs_mq_lit <- cbind(as.data.frame(res_mq_lit2)[,c(1,2,6)],taxa2[rownames(res_mq_lit2),-c(1:3)])
da_asvs_mq_lit[,c(1:8)]

seqtab.r_p_da_mq_lit <- seqtab.r_p[rownames(da_asvs_mq_lit),]
seqtab.r_p_da_mq_lit <- seqtab.r_p_da_mq_lit[,grepl("MQ", colnames(seqtab.r_p_da_mq_lit))]
da_perc_mq_lit <- reshape2::melt(seqtab.r_p_da_mq_lit, 
                                 varnames = c("ASV_ID", "ID"), 
                                 value.name = "Percentage")

da_perc_mq_lit <- merge.data.frame(da_perc_mq_lit, meta, by = "ID")
da_perc_mq_lit$ASV_Species <- paste0(da_perc_mq_lit$ASV_ID, "_", da_asvs_mq_lit$Species)
da_perc_mq_lit$ASV_Species <- factor(da_perc_mq_lit$ASV_Species, levels = rev(unique(da_perc_mq_lit$ASV_Species)))

#da_perc 
asvs_sorted_mq_lit <- rownames(da_asvs_mq_lit)[order(da_asvs_mq_lit$log2FoldChange, decreasing = TRUE)]
asvs_sorted_mq_lit2 <- paste0(asvs_sorted_mq_lit, "_", da_asvs_mq_lit$Species[order(da_asvs_mq_lit$log2FoldChange, decreasing = TRUE)])
da_perc_mq_lit$ASV_Species <- factor(da_perc_mq_lit$ASV_Species, levels = asvs_sorted_mq_lit2)

ggplot(da_perc_mq_lit, aes(x = ASV_Species, y = Percentage, fill = Litter))+
  geom_boxplot()+
  theme_minimal()+
  theme(axis.text.x = element_text(angle = 90))+
  coord_flip()+
  scale_y_continuous(trans='sqrt')
```

```{r}
#Table for paper
da_asvs_mq_lit_cl <- da_asvs_mq_lit[,c(4:9,13,2,3,1)]
da_asvs_mq_lit_cl <- da_asvs_mq_lit_cl[order(da_asvs_mq_lit_cl$log2FoldChange),]
#Correct for the output method
da_asvs_mq_lit_cl$log2FoldChange <- da_asvs_mq_lit_cl$log2FoldChange*-1

```


## MQ Litter vs Sucrose Litter

```{r}
####MQ Litter vs Sucrose litter
res_mql_sucl <- results(dds_c2, contrast = c("condition2", "MQ.Litter", "Sucrose.Litter"))

res_mql_sucl2 <- filterDE(res_mql_sucl)

da_asvs_mql_sucl <- cbind(as.data.frame(res_mql_sucl2)[,c(1,2,6)],taxa2[rownames(res_mql_sucl2),-c(1:3)])
da_asvs_mql_sucl[,c(1:8)]

seqtab.r_p_da_mql_sucl <- seqtab.r_p[rownames(da_asvs_mql_sucl),]
seqtab.r_p_da_mql_sucl <- seqtab.r_p_da_mql_sucl[,grepl("L", colnames(seqtab.r_p_da_mql_sucl))]
da_perc_mql_sucl <- reshape2::melt(seqtab.r_p_da_mql_sucl, 
                                   varnames = c("ASV_ID", "ID"), 
                                   value.name = "Percentage")

da_perc_mql_sucl <- merge.data.frame(da_perc_mql_sucl, meta, by = "ID")
da_perc_mql_sucl$ASV_Species <- paste0(da_perc_mql_sucl$ASV_ID, "_", da_asvs_mql_sucl$Species)
da_perc_mql_sucl$ASV_Species <- factor(da_perc_mql_sucl$ASV_Species, levels = rev(unique(da_perc_mql_sucl$ASV_Species)))

#da_perc 
asvs_sorted_mql_sucl <- rownames(da_asvs_mql_sucl)[order(da_asvs_mql_sucl$log2FoldChange, decreasing = TRUE)]
asvs_sorted_mql_sucl2 <- paste0(asvs_sorted_mql_sucl, "_", da_asvs_mql_sucl$Species[order(da_asvs_mql_sucl$log2FoldChange, decreasing = TRUE)])
da_perc_mql_sucl$ASV_Species <- factor(da_perc_mql_sucl$ASV_Species, levels = asvs_sorted_mql_sucl2)

ggplot(da_perc_mql_sucl, aes(x = ASV_Species, y = Percentage, fill = perfusate))+
  geom_boxplot()+
  theme_minimal()+
  theme(axis.text.x = element_text(angle = 90))+
  coord_flip()+
  scale_y_continuous(trans='sqrt')
```

```{r}
#Table for paper
da_asvs_mql_sucl_cl <- da_asvs_mql_sucl[,c(4:9,13,2,3,1)]
da_asvs_mql_sucl_cl <- da_asvs_mql_sucl_cl[order(da_asvs_mql_sucl_cl$log2FoldChange),]
#Correct for the output method
da_asvs_mql_sucl_cl$log2FoldChange <- da_asvs_mql_sucl_cl$log2FoldChange*-1
```





